Efficient explicit solvers for multipatch discontinuous Galerkin isogeometric analysis

### 2018-06-20 talks/ECCM2018/iga/main.tex
updating ECCM talks
removing old okl, adding okl kernel building

\documentclass[compress]{beamer}
\mode<presentation>
\usetheme{Warsaw}
\usecolortheme{seagull}
%\useoutertheme[subsection=false]{smoothbars}

@@

\newcommand{\uh}{\widehat{u}}
\newcommand{\fnh}{\widehat{f}_n}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\pO}{\partial\Omega}
\newcommand{\Gh}{\Gamma_h}
\newcommand{\Gm}{\Gamma_{-}}
\newcommand{\Gp}{\Gamma_{+}}
\newcommand{\Go}{\Gamma_0}
\newcommand{\Oh}{\Omega_h}

@@

\title[Time-domain IGA]{Time-domain multi-patch\\discontinuous Galerkin methods }
\date[May 19, 2017]{USNCCM 2017\\July 18, 2017}
\author[Chan, Evans]{Jesse Chan\inst{1}, J. A. Evans\inst{2}}
\institute[Rice CAAM]{\inst{1}Department of Computational and Applied Math, Rice University\\ \inst{2} Ann and H.J. Smead Aerospace Engineering Sciences, UC Boulder}

@@

\frame{
\frametitle{High order Galerkin methods for hyperbolic PDEs}
\setcounter{subfigure}{0}
\vspace{-1em}
\begin{columns}
\begin{column}{.5\textwidth}
\begin{itemize}
\item<1-> Time-dependent solutions of hyperbolic equations. 
\vspace{.5em}
\item<2-> Low numerical dissipation and dispersion.
\vspace{.5em}
\item<4-> High order approximations: more accurate per unknown.
\vspace{.5em}
\item<5-> High performance on many-core (explicit time-stepping).
\end{itemize}
\end{column}
\begin{column}{.475\textwidth}
\begin{figure}
\centering
\begin{overlayarea}{\textwidth}{.5\textheight}
\only<1>{
\vspace{-2em}\includegraphics[width=.95\textwidth]{figs/wave.png}
\caption*{Figure courtesy of Axel Modave.}
}
%\only<2>{
%\vspace{1em}
%\includegraphics[width=.95\textwidth]{figs/turbulent1.png}\\
%\vspace{1em}
%\includegraphics[width=.95\textwidth]{figs/turbulent2.png}
%\caption*{Figure courtesy of Per-Olof Persson.}
%}
\only<2>{
\includegraphics[width=.95\textwidth]{figs/wave_N1.eps}
\caption*{\textbf{Fine} linear approximation.}
}
\only<3-4>{
\includegraphics[width=.95\textwidth]{figs/wave_N2.eps}
\caption*{\textbf{Coarse} quadratic approximation.}
}
\only<5>{
\includegraphics[width=.975\textwidth]{figs/gpu.pdf}
\caption*{A graphics processing unit (GPU).}
}
%\caption*{Image courtesy of Axel Modave.}
\end{overlayarea}
\end{figure}
\end{column}
\end{columns}
}

\frame{
\frametitle{Multi-patch discontinuous Galerkin formulations}
\begin{columns}
\begin{column}{.5\textwidth}
\begin{itemize}
\item Model problem: (second order) acoustic wave equation 
\begin{align*}
%\only<1>{
\frac{1}{c^2}\pdd{p}{t}{} - \Delta p &= f.
%}
%\only<2>{
%\pd{v}{t}{} &= \sum_{D^k}\LRp{\Grad p,\grad w}_{D^k} + \LRa{\jump{p},w}\\
%\pd{p}{t}{} &= v.
%}
\end{align*}
\item Pressure-velocity system (first order formulation)
\begin{align*}
\frac{1}{c^2}\pd{p}{t}{} + \Div \bm{u} &= f\\
\pd{\bm{u}}{t}{} + \Grad p &= 0.
\end{align*}
\end{itemize}
\end{column}

@@

\vspace{1em}
\begin{itemize}
\item Weak patch coupling through DG numerical flux (SIPG, upwind).  
\end{itemize}

\let\thefootnote\relax\footnotetext{\tiny Langer et al 2014. Multipatch discontinuous Galerkin isogeometric analysis.}
\let\thefootnote\relax\footnotetext{\tiny Wilcox et al 2010. A high-order DG method for wave propagation through coupled elastic-acoustic media.}
}

\frame{
\frametitle{Method of lines multi-patch DG discretization}
\vspace{-1em}
\begin{itemize}
\item Semi-discrete system:
\[
\bm{M}_h\td{\bm{u}}{t} = \bm{A}_h\bm{u} \quad \Rightarrow \quad \td{\bm{u}}{t} = \bm{M}_h^{-1}\bm{A}_h\bm{u}.
\]
\item Global mass matrix $\bm{M}_h$ is (patch) block diagonal.  
\vspace{1em}
\item Tailored patch discretizations: explicit time-stepping, curved domains. 
%\vspace{.25em}
\begin{itemize}
\item Efficient mass matrix inversion over each patch.
\vspace{.5em}
\item High order accuracy, energy stability on curved geometries.
\vspace{.5em}
\item Improved stable timestep restriction (CFL).
\end{itemize}
\end{itemize}
}

\section{Spline spaces and optimal knot vectors }
\frame[noframenumbering]{
\frametitle{Outline}
\only<1>{
\tableofcontents
}
\only<2>{
\tableofcontents[currentsection]
}
}

%\frame{
%\frametitle{Spline spaces and sup-infs}
%
%\begin{itemize}
%\item Let $X$ be Banach, and $X_n$ be some $n$-dimensional subspace of $X$
%\item 
%\item The $n$-width of $X$ is then defined to be the infimum over all subspaces $X_n$ with ${\rm dim}\LRp{X_n} = n$, and a specific subspace $X_n$ is referred to as an \textit{optimal} subspace if its sup-inf attains the $n$-width.
%\item \note{FINISH}
%\end{itemize}
%}
\frame{
\frametitle{B-spline basics}

\begin{itemize}
\item Standard 1D B-splines: $B^0_i(x) = \mathbbm{1}_{\xi_i \leq x \leq \xi_{i+1}}$, 
\[
B^k_i(x) = \frac{x - \xi_i}{\xi_{i+p}-\xi_i}B^{k-1}_i(x) + \frac{\xi_{i+p+1}-x}{\xi_{i+p+1}-\xi_{i+1}}B^{k-1}_{i+1}(x).
\]
\item Assume patches are mapped quads/cubes: tensor product basis
\[
B^p_{ijk}(x,y,z) = B^p_{i}(x)B^p_{j}(y)B^p_{k}(z).
\]
\item Assume maximally continuous, open knot vectors
\begin{align*}
%\bm{\Xi} &= \LRc{-1 = \xi_1,\ldots,\xi_{{2p}+K+1} = 1}, \qquad \xi_i \leq \xi_{i+1},\\
\xi_{p+1} &< \ldots < \xi_{p+1+K},\\
\xi_1 &= \ldots = \xi_{p+1},\\
\xi_{p+1 + K} &= \ldots = \xi_{2p+1+K}.\\
\end{align*}
\end{itemize}
}
\frame{
\frametitle{B-spline bases and optimal spline spaces}
\vspace{-1em}
\begin{figure}
\centering
\subfloat[Uniform knots]{\includegraphics[width=.32\textwidth]{figs/unifknots.png}}
\hspace{2.5em}
\subfloat[Optimal knots]{\includegraphics[width=.32\textwidth]{figs/optknots.png}}
%\caption{B-spline bases with uniform and optimal knot vectors ($p = 4, K = 4$).}
\end{figure}

>>>>

\documentclass[compress]{beamer}
\mode<presentation>
\usetheme{Warsaw}
\usecolortheme{seagull}
%\useoutertheme[subsection=false]{smoothbars}
\usepackage{fancybox}

@@

\newcommand{\uh}{\widehat{u}}
\newcommand{\fnh}{\widehat{f}_n}
\renewcommand{\hat}{\widehat}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\pO}{\partial\Omega}
\newcommand{\Gh}{\Gamma_h}
\newcommand{\Gm}{\Gamma_{-}}
\newcommand{\Gp}{\Gamma_{+}}
\newcommand{\Go}{\Gamma_0}
\newcommand{\Oh}{\Omega_h}

@@

\title[Multi-patch IGA]{Efficient explicit solvers for multipatch discontinuous Galerkin isogeometric analysis}
\date[June 9-15, 2018]{ECCM-ECFD 2018\\June 9-15, 2018}
\author[Chan, Evans]{Jesse Chan\inst{1}, J. A. Evans\inst{2}}
\institute[Rice CAAM]{\inst{1}Department of Computational and Applied Math, Rice University\\ \inst{2} Ann and H.J. Smead Aerospace Engineering Sciences, UC Boulder}

@@

\frame{
\frametitle{High order methods for hyperbolic PDEs}
\setcounter{subfigure}{0}
\vspace{-1em}
\begin{columns}
\begin{column}{.5\textwidth}
\begin{itemize}
\item<1-> Time-dependent solutions of hyperbolic equations. 
\vspace{.5em}
\item<2-> Low numerical dissipation and dispersion.
\vspace{.5em}
\item<4-> High order approximations: more accurate per unknown.
\vspace{.5em}
\item<5-> High performance on many-core (explicit time-stepping).
\end{itemize}
\end{column}
\begin{column}{.475\textwidth}
\begin{figure}
\centering
\begin{overlayarea}{\textwidth}{.5\textheight}
\only<1>{
\vspace{-2em}\includegraphics[width=.95\textwidth]{figs/wave.png}
\caption*{Figure courtesy of Axel Modave.}
}
%\only<2>{
%\vspace{1em}
%\includegraphics[width=.95\textwidth]{figs/turbulent1.png}\\
%\vspace{1em}
%\includegraphics[width=.95\textwidth]{figs/turbulent2.png}
%\caption*{Figure courtesy of Per-Olof Persson.}
%}
\only<2>{
\includegraphics[width=.95\textwidth]{figs/wave_N1.eps}
\caption*{\textbf{Fine} linear approximation.}
}
\only<3-4>{
\includegraphics[width=.95\textwidth]{figs/wave_N2.eps}
\caption*{\textbf{Coarse} quadratic approximation.}
}
\only<5>{
\includegraphics[width=.975\textwidth]{figs/gpu.pdf}
\caption*{A graphics processing unit (GPU).}
}
%\caption*{Image courtesy of Axel Modave.}
\end{overlayarea}
\end{figure}
\end{column}
\end{columns}
}

\frame{
\frametitle{Multi-patch discontinuous Galerkin formulations}
\begin{columns}
\begin{column}{.5\textwidth}
\begin{itemize}
\item Wave propagation problems (acoustics, Maxwells, elasticity).
\vspace{.1em}
%\begin{align*}
%%\only<1>{
%\frac{1}{c^2}\pdd{p}{t}{} - \Delta p &= f.
%%}
%%\only<2>{
%%\pd{v}{t}{} &= \sum_{D^k}\LRp{\Grad p,\grad w}_{D^k} + \LRa{\jump{p},w}\\
%%\pd{p}{t}{} &= v.
%%}
%\end{align*}
\item Model problem: acoustic wave equation (both first and second order formulations)
\begin{align*}
\frac{1}{c^2}\pd{p}{t}{} + \Div \bm{u} &= f\\
\pd{\bm{u}}{t}{} + \Grad p &= 0.
\end{align*}
%\item Results extend to second order formulation. 
%\item Multiple geometric patches,\\IGA on each patch.  
\end{itemize}
\end{column}

@@

\vspace{1em}
\begin{itemize}
\item Multiple geometric patches, weak patch coupling through DG-like numerical interface flux (SIPG, upwind).  
\end{itemize}

\let\thefootnote\relax\footnotetext{\tiny Langer et al (2014). \textit{Multipatch discontinuous Galerkin isogeometric analysis}.}
\let\thefootnote\relax\footnotetext{\tiny Wilcox et al (2010). \textit{A high-order DG method for wave propagation through coupled elastic-acoustic media}.}
}

\frame{
\frametitle{Explicit solvers for multi-patch DG-IGA}
%\vspace{-1em}
\begin{itemize}
\item Semi-discrete system:
\[
\bm{M}_h\td{\bm{u}}{t} = \bm{A}_h\bm{u} \quad \Rightarrow \quad \td{\bm{u}}{t} = \bm{M}_h^{-1}\bm{A}_h\bm{u}.
\]
\vspace{.125em}
\item Global mass matrix $\bm{M}_h$ is (patch) block diagonal.  
\vspace{1.5em}
\item Challenges and questions: 
\vspace{.5em}
\begin{itemize}
\item<2-> How to efficiently invert 3D patch mass matrices while guaranteeing accuracy and \textcolor{red}{stability} (especially for explicit methods)?  
\vspace{.5em}
\item<3-> Do splines/IGA offer advantages over $C^0$-FEM/DG for explicit solvers?  
\vspace{.5em}
\item<4-> Can we tailor IGA discretizations towards explicit time-stepping?
\end{itemize}

%\vspace{1em}
%\item Explicit time-stepping, curved domains. 
%%\vspace{.25em}
%\begin{itemize}
%\item Efficient mass matrix inversion over each patch.
%\vspace{.5em}
%\item High order accuracy, energy stability on curved geometries.
%\vspace{.5em}
%\item Improved stable timestep restriction (CFL).
%\end{itemize}
\end{itemize}
}

\section{Inverting IGA mass matrices }
\frame[noframenumbering]{
\frametitle{Outline}
\only<1>{
\tableofcontents
}
\only<2>{
\tableofcontents[currentsection]
}
}

\frame{
\frametitle{B-splines: assumptions for this talk}

\begin{itemize}
\item Reference coordinates $\hat{\bm{x}} \in \hat{D}$, physical coordinates $\bm{x} \in D^k$.  
\vspace{.5em}
\item Standard 1D B-splines: $B^0_i(\hat{x}) = \mathbbm{1}_{\xi_i \leq \hat{x} \leq \xi_{i+1}}$, 
\[
B^k_i(\hat{x}) = \frac{\hat{x} - \xi_i}{\xi_{i+p}-\xi_i}B^{k-1}_i(\hat{x}) + \frac{\xi_{i+p+1}-\hat{x}}{\xi_{i+p+1}-\xi_{i+1}}B^{k-1}_{i+1}(\hat{x}).
\]
\item Physical basis: mapping of tensor product basis on reference domain
%\begin{overlayarea}{\textwidth}{.17\textheight}
\[
%\only<1>{
B^p_{ijk}(\hat{\bm{x}}) = B^p_{i}(\hat{x})B^p_{j}(\hat{y})B^p_{k}(\hat{z}).%, \qquad \text{(mapped B-splines)}}
%\only<2>{B^p_{ijk}({\bm{x}}) = \frac{B^p_{i}(\hat{x})B^p_{j}(\hat{y})B^p_{k}(\hat{z})}{w_R(\hat{\bm{x}})}, \qquad \text{(NURBS)}}
\]
%\end{overlayarea}
\item Assume maximally continuous, open knot vectors
\begin{align*}
%\bm{\Xi} &= \LRc{-1 = \xi_1,\ldots,\xi_{{2p}+K+1} = 1}, \qquad \xi_i \leq \xi_{i+1},\\
\xi_{p+1} &< \ldots < \xi_{p+1+K},\\
\xi_1 &= \ldots = \xi_{p+1},\\
\xi_{p+1 + K} &= \ldots = \xi_{2p+1+K}.\\
\end{align*}
\end{itemize}
}

\frame{
\frametitle{IGA mass matrices: stability/accuracy vs efficiency}

\begin{itemize}
\item Energy stability: if \note{$\bm{u}^T\bm{A}_h\bm{u} \leq 0$}, semi-discrete solution won't blow up
\[
\bm{M}_h\td{\bm{u}}{t} = \bm{A}_h\bm{u} \quad \Longrightarrow \quad \td{}{t}\LRp{\bm{u}^T\bm{M}_h \bm{u}} =  \frac{1}{2}\pd{}{t}{}\nor{u}_{L^2}^2 \leq 0.
\]
\item Approximating $\bm{M}_h^{-1}$ impacts semi-discrete \textcolor{red}{stability} and \textcolor{red}{accuracy}.  
\vspace{2em}
\item<2-> Curved patch mass matrices $\bm{M}_J$: tensor product basis $B^p_{ijk}(\hat{\bm{x}})$
\begin{overlayarea}{\textwidth}{.185\textheight}
\[
\only<1-2>{\LRp{\bm{M}_J}_{ijk,lmn} = \int_{\widehat{D}} B^p_{ijk}(\widehat{\bm{x}})B^p_{lmn}(\widehat{\bm{x}}) \textcolor{red}{J(\widehat{\bm{x}})}\diff{\widehat{\bm{x}}}, \qquad \text{(B-splines)}}
\only<3>{\LRp{\bm{M}_J}_{ijk,lmn} = \int_{\widehat{D}} \frac{B^p_{ijk}(\widehat{\bm{x}})}{\textcolor{red}{w_R\LRp{\hat{\bm{x}}}}}\frac{B^p_{lmn}(\widehat{\bm{x}})}{\textcolor{red}{w_R\LRp{\hat{\bm{x}}}}} \textcolor{red}{{J(\widehat{\bm{x}})}}\diff{\widehat{\bm{x}}} \qquad \text{(NURBS)}}
\only<4>{\LRp{\bm{M}_J}_{ijk,lmn} = \int_{\widehat{D}} B^p_{ijk}(\widehat{\bm{x}})B^p_{lmn}(\widehat{\bm{x}}) \textcolor{red}{\frac{J(\widehat{\bm{x}})}{w^2_R\LRp{\hat{\bm{x}}}}}\diff{\widehat{\bm{x}}} \qquad \text{(NURBS)}}
\]
\end{overlayarea}
\item<2-> No Kronecker structure due to $J(\widehat{\bm{x}})$: $\bm{M}_h$ \textcolor{red}{expensive} to invert in 3D!

\end{itemize}
}

\begin{itemize}
\item Mass lumping: loss of high order accuracy for IGA.
\vspace{.75em}
\item Preconditioning: 
\vspace{.35em}
\begin{itemize}
\item Additional cost and complexity for a time-domain code.
\vspace{.35em}
\item Semi-discrete stability requires approximation of $\bm{M}_J^{-1}$ to induce a norm on $\bm{u}$ (e.g.\ a fixed symmetric positive-definite linear operator).
\vspace{.35em}
\item Example: Krylov methods approximate $\bm{M}_J^{-1}$ as a non-linear operator!
\end{itemize}
\vspace{.75em}
\item Isogeometric collocation: restores tensor product structure, but semi-discrete stability is more difficult to prove.  
\end{itemize}

\let\thefootnote\relax\footnotetext{\tiny Gao and Calo (2014). \textit{Fast isogeometric solvers for explicit dynamics}.}
\let\thefootnote\relax\footnotetext{\tiny Evans, Hiemstra, Hughes, Reali (2017).  \textit{Explicit higher-order accurate IG collocation methods for structural dynamics}.}
\let\thefootnote\relax\footnotetext{\tiny Wathen and Rees (2009). \textit{Chebyshev semi-iteration in preconditioning for problems including the mass matrix}.}
\let\thefootnote\relax\footnotetext{\tiny Auricchio et al (2012). \textit{Isogeometric collocation for elastostatics and explicit dynamics}.}
}

\frame{
\frametitle{An energy stable and efficient approximation to $\bm{M}^{-1}$}

\begin{itemize}
\item Replace $\bm{M}_J$ with symmetric pos-def ``weight-adjusted'' approximation:
\[
%\td{}{t}
 \bm{M}_J\bm{u} \Rightarrow %\td{}{t}
 \textcolor{red}{\widehat{\bm{M}}\bm{M}_{1/J}^{-1}\widehat{\bm{M}}} \bm{u}, \qquad \LRp{\widehat{\bm{M}}}_{ijk,lmn} = \int_{\widehat{D}} B^p_{ijk}(\widehat{\bm{x}})B^p_{lmn}(\widehat{\bm{x}})\diff{\widehat{\bm{x}}}.
\]
%\vspace{.5em}
\item Weight-adjusted inverse: Kronecker product, matrix-free eval.\ of $\bm{M}_{1/J}$
\begin{align*}
\bm{M}^{-1}_J &\approx \LRp{\widehat{\bm{M}}\bm{M}_{1/J}^{-1}\widehat{\bm{M}}}^{-1}  = \widehat{\bm{M}}^{-1} \bm{M}_{1/J} \widehat{\bm{M}}^{-1}\\
\widehat{\bm{M}}^{-1} &= \widehat{\bm{M}}^{-1}_{\rm 1D} \otimes \widehat{\bm{M}}^{-1}_{\rm 1D} \otimes \widehat{\bm{M}}^{-1}_{\rm 1D}.
\end{align*}
\item Energy stability with respect to an \note{equivalent} norm 
\[
C_1(J) \nor{\bm{u}}_{\widehat{\bm{M}}\bm{M}_{1/J}^{-1}\widehat{\bm{M}}} \leq \nor{\bm{u}}_{\bm{M}_J} \leq C_2 \nor{\bm{u}}_{\widehat{\bm{M}}\bm{M}_{1/J}^{-1}\widehat{\bm{M}}}.
\]
\end{itemize}

\let\thefootnote\relax\footnotetext{\tiny Chan, Hewett, Warburton (2016).  \textit{Weight-adjusted DG methods: wave prop.\ in heterogeneous media}. }
\let\thefootnote\relax\footnotetext{\tiny Chan, Hewett, Warburton (2016).  \textit{Weight-adjusted DG methods: curvilinear meshes}. }
}

\frame{
\frametitle{Accuracy: weighted vs weight-adjusted mass matrix}

\begin{figure}
\centering
\only<1>{
\subfloat[Warped mesh]{\includegraphics[width=.425\textwidth]{figs/mapped.png}}
\subfloat[Error for acoustics (first order form)]{
\begin{tikzpicture}
\begin{loglogaxis}[
    legend cell align=left,
    legend style={legend pos=south east, font=\tiny},
    width=.5\textwidth,
     ymin=1e-17, ymax=2,    
    xlabel={Mesh size $h$},
    ylabel={$L^2$ error}, 
    grid style=dashed,
] 
\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.154255)(0.25,0.0280139)(0.125,0.000775805)(0.0625,7.51259e-06)};
\addplot[color=red,mark=x,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.154523)(0.25,0.0280342)(0.125,0.000775107)(0.0625,7.51217e-06)};
\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,7.56052e-06)(0.25,1.32834e-07)(0.125,1.51285e-11)(0.0625,1.69165e-17)};

\legend{$L^2$ projection,Weight-adjusted,Difference}
%\legend{Uniform, Optimal, Smoothed}
\end{loglogaxis}
\end{tikzpicture}
}
%\subfloat[$L^2$ errors ]{
%\begin{tikzpicture}
%\begin{loglogaxis}[
%    legend cell align=left,
%    legend style={legend pos=south east, font=\tiny},
%    width=.5\textwidth,
%    xlabel={Mesh size $h$},
%    ylabel={$L^2$ error}, 
%             ymin=1e-10, ymax=2,    
%    grid style=dashed,
%] 
%\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
%coordinates{(0.5,0.0110908)(0.25,0.00138298)(0.125,2.29018e-05)(0.0625,3.18561e-07)};
%\addplot[color=red,mark=x,semithick, mark options={solid,fill=markercolor}]
%coordinates{(0.5,0.0111346)(0.25,0.00138582)(0.125,2.30061e-05)(0.0625,3.18565e-07)};
%\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
%coordinates{(0.5,0.000986363)(0.25,8.87776e-05)(0.125,2.18816e-06)(0.0625,1.65013e-09)};
%\legend{$L^2$ projection,Weight-adjusted,Difference}
%%\legend{Uniform, Optimal, Smoothed}
%\end{loglogaxis}
%\end{tikzpicture}}
}
\only<2>{
\subfloat[Warped mesh]{\includegraphics[width=.425\textwidth]{figs/mapped2.png}}
\subfloat[Error for acoustics (first order form)]{
\begin{tikzpicture}
\begin{loglogaxis}[
    legend cell align=left,
    legend style={legend pos=south east, font=\tiny},
    width=.5\textwidth,
     ymin=1e-17, ymax=2,    
    xlabel={Mesh size $h$},
    ylabel={$L^2$ error}, 
    grid style=dashed,
] 
\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.338603)(0.25,0.143169)(0.125,0.0234825)(0.0625,0.00033695)};
\addplot[color=red,mark=x,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.388935)(0.25,0.146233)(0.125,0.0235623)(0.0625,0.000336824)};
\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.00977253)(0.25,0.000187168)(0.125,3.16763e-07)(0.0625,9.44863e-12)};

\legend{$L^2$ projection,Weight-adjusted,Difference}
%\legend{Uniform, Optimal, Smoothed}
\end{loglogaxis}
\end{tikzpicture}
}
%\subfloat[$L^2$ errors ]{
%\begin{tikzpicture}
%\begin{loglogaxis}[
%    legend cell align=left,
%    legend style={legend pos=south east, font=\tiny},
%    width=.5\textwidth,
%     ymin=1e-10, ymax=2,    
%    xlabel={Mesh size $h$},
%    ylabel={$L^2$ error}, 
%    grid style=dashed,
%] 
%\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
%coordinates{(0.5,0.0444376)(0.25,0.0144725)(0.125,0.000702933)(0.0625,6.79061e-06)};
%\addplot[color=red,mark=x,semithick, mark options={solid,fill=markercolor}]
%coordinates{(0.5,0.1182)(0.25,0.017419)(0.125,0.000745669)(0.0625,6.79513e-06)};
%\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
%coordinates{(0.5,0.109529)(0.25,0.00969371)(0.125,0.000248811)(0.0625,2.47881e-07)};
%\legend{$L^2$ projection,Weight-adjusted,Difference}
%%\legend{Uniform, Optimal, Smoothed}
%\end{loglogaxis}
%\end{tikzpicture}}
}
\caption{$L^2$ errors for the acoustic wave equation using weighted and weight-adjusted mass matrices for tensor product $p = 4$ splines. }

\frame{
\frametitle{Accuracy of the weight-adjusted mass matrix}

\begin{itemize}
\item Difference between the weighted $L^2$ and weight-adjusted inner products is high order accurate: for $v(\bm{x})$ of degree $q$, 
\begin{align*}
&\LRb{ \bm{v}^T\bm{M}_J\bm{u} - \bm{v}^T\widehat{\bm{M}}\bm{M}_{1/J}^{-1}\widehat{\bm{M}}\bm{u}}\\
&\leq C_{J} \nor{J}_{W^{p+1,\infty}\LRp{D^k}} h^{2p+2-q} \nor{u}_{W^{p+1,2}\LRp{D^k}}.
\end{align*}
\vspace{.5em}
\item Difference between $L^2$ and weight-adjusted projection is \textcolor{red}{$O(h^{p+2})$}!
\begin{align*}
\nor{P_h u - \tilde{P}_h u}_{L^2\LRp{D^k}} &\lesssim \nor{\frac{1}{\sqrt{J}}}_{L^{\infty}}^2 \nor{J}_{W^{p+1,\infty}\LRp{D^k}} h^{p+2}\nor{u}_{W^{p+1,2}\LRp{D^k}}.
\end{align*}
\end{itemize}

\let\thefootnote\relax\footnotetext{\tiny Chan, Wilcox (2018). \textit{On discretely entropy stable weight-adjusted discontinuous Galerkin methods: curvilinear meshes}.}
}

\section{Why splines vs $C^0$-FEM?  CFL restrictions}

\frame[noframenumbering]{
\tableofcontents[currentsection]
}

\frame{
\frametitle{Estimating the CFL restriction}

\begin{itemize}
%\item Maximum stable timestep (CFL) depends on $h, p$, and wavespeed.
\item For explicit time-stepping method: estimate $dt \propto \frac{1}{\max\LRb{\lambda_j}}$
\[
\bm{M}_h \bm{v} = \lambda \bm{A}_h\bm{v}.
\]
\item Bendixon-Hirsch lemma: bound ${\rm Re}\LRp{\lambda_j}, {\rm Im}\LRp{\lambda_j}$ using the symmetric and skew-symmetric parts  of $\bm{A}_h$
\[
\underbrace{\frac{1}{2}\LRp{\bm{A}_h+\bm{A}_h^T}}_{\bm{A}_{\rm sym}}+ \underbrace{\frac{1}{2}\LRp{\bm{A}_h-\bm{A}_h^T}}_{\bm{A}_{\rm skew}}, \qquad
\begin{array}{c}
\LRb{{\rm Re}\LRp{\lambda_j}} \leq \rho\LRp{\bm{M}^{-1}_h\bm{A}_{\rm sym}},\\
\LRb{{\rm Im}\LRp{\lambda_j}} \leq \rho\LRp{\bm{M}^{-1}_h\bm{A}_{\rm skew}}.
\end{array}
\]
\item $\rho\LRp{\bm{M}_h^{-1}\bm{A}_{\rm sym}}, \rho\LRp{\bm{M}_h^{-1}\bm{A}_{\rm skew}}$: generalized Rayleigh quotients
\[
\rho\LRp{\bm{M}_h^{-1}\bm{A}_{\rm sym}} = \frac{\bm{u}^T\bm{A}_{\rm sym}\bm{u}}{\bm{u}^T\bm{M}_h\bm{u}}, \qquad 
\rho\LRp{\bm{M}_h^{-1}\bm{A}_{\rm skew}} = \frac{\LRb{\bm{u}^* (i\bm{A}_{\rm skew})\bm{u}}}{\bm{u}^*\bm{M}_h\bm{u}}
\]
\end{itemize}
}

\frame{
\frametitle{CFL: constants in trace and inverse inequalities}
\begin{itemize}
\item For hyperbolic problems (advection, acoustics), can bound
\begin{align*}
&\frac{\bm{u}^T\bm{A}_{\rm sym}\bm{u}}{\bm{u}^T\bm{M}_h\bm{u}} \lesssim \frac{\nor{u}_{L^2\LRp{\partial D^k}}^2}{\nor{u}_{L^2\LRp{D^k}}^2} \lesssim \frac{C_T}{h}, \\
&\frac{\LRb{\bm{u}^* (i\bm{A}_{\rm skew})\bm{u}}}{\bm{u}^*\bm{M}_h\bm{u}} \lesssim \frac{\nor{\Grad u}_{L^2\LRp{D^k}}}{\nor{u}_{L^2\LRp{D^k}}} \lesssim \frac{C_I}{h}.
\end{align*}
\item $C_T, C_I$: \note{$p$-dependent constants} in trace, inverse inequalities
\[
\nor{\Grad u}_{L^2\LRp{\hat{D}}} \leq C_I \nor{u}_{L^2\LRp{\hat{D}}} , \qquad \nor{u}_{L^2\LRp{\partial \hat{D}}}^2 \leq C_T \nor{u}_{L^2\LRp{\hat{D}}}^2.
\]
\item Summary: $dt \propto \frac{1}{\max\LRb{\lambda_j}} \leq \frac{h}{\max{C_T, C_I}}$.  What do $C_T, C_I$ look like?  Can compute $C_T, C_I$ using a generalized eigenvalue problem.    %For $C^0$-FEM and DG, $dt \leq O\LRp{\frac{h}{p^2}}$.
\end{itemize}
}

\frame{
\frametitle{Trace and inverse inequality constants: $C^0$-FEM vs splines}
\vspace{-.5em}
\begin{figure}
\centering
\only<1>{\includegraphics[width=\textwidth]{figs/Definitions.pdf}
\caption{A parametric patch has $K$ elements per side, while a physical patch has size $H$.  The mesh resolution is $h = H/K$.}}
\only<2->{
\setcounter{subfigure}{0}
\subfloat[Inverse constants, $K = 2p$]{
\begin{tikzpicture}
\begin{axis}[
width=.45\textwidth,
    xlabel={Degree of approximation $p$},   
%    ylabel={$C_I / K$},
    xmin=.5, xmax=12.5,
    ymax=0,ymax=90,    
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
    xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
coordinates{(1,2)(2,4.5)(3,8)(4,12.5)(5,18)(6,24.5)(7,32)(8,40.5)(9,50)(10,60.5)(11,72)(12,84.5)};
\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
coordinates{(1,2)(2,3.42857)(3,5.27022)(4,7.3491)(5,9.61122)(6,12.0258)(7,14.5716)(8,17.233)(9,19.9977)(10,22.8559)(11,25.7995)(12,28.8217)};
%\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
%coordinates{(1,2)(2,3.15356)(3,4.41024)(4,5.72791)(5,7.09105)(6,8.48919)(7,9.91299)(8,11.3583)(9,12.8221)(10,14.3019)(11,15.796)(12,17.3028)};

\legend{Polynomial $C_I$, Spline $C_I/K$}%, Smoothed $C_I/K$ }
\end{axis}
\end{tikzpicture}
}
\hspace{1em}
\subfloat[Trace constants, $K = 2p$]{
\begin{tikzpicture}
\begin{axis}[
width=.45\textwidth,
    xlabel={Degree of approximation $p$},   
    xmin=.5, xmax=12.5,
    ymax=0,ymax=65,    
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
    xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
coordinates{(1,1.73205)(2,3.87298)(3,6.5216)(4,9.74981)(5,13.5914)(6,18.0596)(7,23.1598)(8,28.894)(9,35.2634)(10,42.2685)(11,49.9096)(12,58.187)};
\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
coordinates{(1,1.73205)(2,2.43668)(3,3.48657)(4,4.75661)(5,6.16327)(6,7.6751)(7,9.27477)(8,10.9506)(9,12.694)(10,14.4983)(11,16.3578)(12,18.2682)};
%\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
%coordinates{(1,1.73205)(2,2.32019)(3,2.9916)(4,3.77273)(5,4.61001)(6,5.4806)(7,6.37331)(8,7.28317)(9,8.20701)(10,9.14265)(11,10.0885)(12,11.0432)};

\legend{Polynomial $C_T$, Spline $C_T/K$}%, Smoothed $C_T/K$ }
\end{axis}
\end{tikzpicture}
}
\caption*{Polynomial constants are $O(p^2)$, observed spline constants \textcolor{red}{$O(p)$} for $K \geq O(1/p)$.}
\let\thefootnote\relax\footnotetext{\tiny Takacs, Takacs (2016). \textit{Approximation error estimates and inverse inequalities for B-splines of maximum smoothness}.}
}
\end{figure}
\only<2->{\uncover<3->{
\begin{center}
\ovalbox{\textcolor{red}{$dt \propto \frac{h}{p^2}$} for $C^0$-FEM and DG, \textcolor{red}{$dt \propto \frac{h}{p}$} for IGA!}
\end{center}}}
}

\section{Optimizing splines for explicit solvers}
\frame[noframenumbering]{
\tableofcontents[currentsection]
}
